---
title: "Homework 2"
author: "Kristi Chau"
date: "9/27/2019"
output: github_document
fig_caption: yes
latex_engine: xelatex
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{booktabs}
   - \usepackage{longtable}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{float}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{tabu}
   - \usepackage{threeparttable}
   - \usepackage{threeparttablex}
   - \usepackage[normalem]{ulem}
   - \usepackage[normalem]{ulem}
   - \usepackage[utf8]{inputenc}
   - \usepackage{makecell}
   - \usepackage{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(knitr)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
```

## Problem 1

Read and clean Mr. Trash Wheel sheet

```{r}
mtw_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "Mr. Trash Wheel!A2:N338") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))

##specified sheet in range
```

Read and clean precipitation data for 2017 and 2018 and combine them

```{r}
precip_2017_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "2017 Precipitation!A2:B14") %>% 
  janitor::clean_names() %>%
  mutate(year = 2017)

precip_2018_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",range = "2018 Precipitation!A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2018)

precip_tidy = 
  bind_rows(precip_2017_df, precip_2018_df) %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[month]) %>%
  rename(total_precip = total) %>%
  select(month, year, total_precip)
```

There are `r nrow(mtw_df)` rows and `r ncol(mtw_df)` columns in the Mr. Trash Wheel dataset, giving us `r nrow(mtw_df) * ncol(mtw_df)` datapoints. Some variables in this dataset are: `r colnames(mtw_df)[c(1,4,5,14)]`. The median number of sports balls collected in dumpsters in 2017 was `r median(pull(mtw_df,sports_balls)[ pull(mtw_df,year) == 2017])`. 

The Precipitation dataset has `r nrow(precip_tidy)` rows and `r ncol(precip_tidy)` columns, which is `r nrow(precip_tidy) * ncol(precip_tidy)` datapoints. The precipitation dataset contains the variables: `r colnames(precip_tidy)`. The total precipitation in 2018 was `r sum(pull(precip_tidy,total_precip)[pull(precip_tidy,year) == 2018])` inches.

## Problem 2

Cleaning the data in pols-month.csv

```{r}
pols_month_df = 
  read.csv("./data/pols-month.csv") %>%
  separate(mon, c("year","month","day"),sep = "-") %>%
  mutate(month = month.name[as.integer(month)], year = as.integer(year), president = ifelse(prez_gop %in% c(1,2), "gop", ifelse(prez_dem %in% c(1,2), "dem", NA))) %>%
  select(-prez_gop, -prez_dem, -day)
```

Cleaning the data in snp.csv

```{r}
snp_df = 
  read.csv("./data/snp.csv") %>%
  separate(date, c("day","month","year"),"/") %>%
  mutate(month = month.name[as.integer(month)], year = as.integer(year)) %>%
  rename(closing_stock_index = close) %>%
  select(year, month, closing_stock_index, -day)
```

Cleaning the unemployment data

```{r}
unemployment_df = 
  read.csv("./data/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_percent"
  ) %>%
  mutate(month = month.name[match(month,month.abb)]) %>%
  janitor::clean_names()
```

Merging the datasets

```{r}
pols_snp_data = 
  left_join(pols_month_df, snp_df, by = c("year","month"))

fte_data = 
  left_join(pols_snp_data, unemployment_df, by = c("year","month"))
```

The `pols_month_df` contained data about the number of national politicians who are democratic or republic at any given time. The `snp_df` contained data about the S&P's stock market index on dates of observation. The `unemployment_df` contained data about the percentage of unemployment at dates of measurement. The resulting merged dataset has `r nrow(fte_data)` rows and `r ncol(fte_data)` columns containing `r nrow(fte_data) * ncol(fte_data)` datapoints. The years range from `r min(pull(fte_data,year))` to `r max(pull(fte_data,year))`. Some key variables are: `r colnames(fte_data)[c(1,2,9,10,11)]`.

## Problem 3

```{r}
baby_names_data =
  read.csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(childs_first_name = str_to_title(child_s_first_name), ethnicity = case_when(ethnicity == "ASIAN AND PACI"~"ASIAN AND PACIFIC ISLANDER", ethnicity == "BLACK NON HISP"~"BLACK NON HISPANIC", ethnicity == "WHITE NON HISP"~"WHITE NON HISPANIC", TRUE~as.character(ethnicity))) %>%
  select(-child_s_first_name) %>%
  distinct()
```

Creating reader-friendly tables

```{r}
baby_names_data %>%
  filter(childs_first_name == "Olivia") %>%
  arrange(year_of_birth) %>%
  select(year_of_birth, rank, ethnicity) %>%
  kable(caption = "A ranking of the baby name Olivia over time")

baby_names_data %>%
  filter(gender == "MALE" & rank == 1) %>%
  arrange(year_of_birth) %>%
  select(year_of_birth, childs_first_name, ethnicity) %>%
  kable(caption = "Most popular male baby names over time")
```

Create a scatter plot

```{r}
pop_vs_freq_names = 
  baby_names_data %>%
  filter(gender == "MALE" & year_of_birth == 2016 & ethnicity == "WHITE NON HISPANIC") %>%
  ggplot(aes(x = rank, y = count)) +
  geom_point() +
  labs(
    title = "Name Popularity vs Frequency of Name for Male, White non-Hispanic Children Born in 2016") + 
    xlab("popularity ranking of a name") + 
    ylab("# of children with a name") + 
    theme(plot.title = element_text(size = 11, hjust = 0.5)
  )

ggsave("pop_vs_freq_names.pdf", pop_vs_freq_names, width = 8, height = 5)

pop_vs_freq_names
```